\TheTitle

### 2019-07-11 Paper1/framework.tex
added adaptive one shot to paper and polished up some parts

\subsection{Costs of the numerical flux}
A straight forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}. By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available. However, the need to approximate the flux by an integral seems to be numerically expensive, especially when considering that the stochastic-Galerkin method often allows an analytic computation of all arising integrals. In the following, we discuss this issue and show the advantages of our numerical flux choice by considering the stochastic Burgers' equation
\begin{align*}
\partial_t u + \partial_x \frac{u^2}{2} &= 0,\\
u(t=0,x,\xi) &= u_{IC}(x,\xi).
\end{align*}
The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials. Testing with these basis functions and choosing the SG closure \eqref{eq:SGClosure} yields the closed SG moment system
\begin{align*}
\partial_t \hat u_i + \partial_x \frac12\sum_{n,m = 0}^M \hat u_n \hat u_m \langle \varphi_n\varphi_m\varphi_i \rangle = 0.
\end{align*}
Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{M\times M}$ gives
\begin{align*}
\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = 0
\end{align*}
with $F_i(\bm{\hat u}) = \frac12\bm{\hat u}^T\bm C_i\bm{\hat u}$. Note that $\bm{C}_i$ can be computed analytically, hence choosing a Lax-Friedrichs flux
\begin{align}\label{eq:numFluxAnalytic}
G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i
\end{align}
requires no integral evaluations. Recall, that the numerical flux choice made in this work gives
\begin{align}\label{eq:numericalFluxIPMBurgers}
 \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \langle g(\mathcal{U}_{SG}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}_{SG}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}^T\rangle^T_Q.
\end{align}
When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the integral brackets is of order $3M$. Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 M +\frac12$ quadrature points suffice for an exact computation of the numerical flux. Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent. 
Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When precomputing the values $\mathcal{U}_{SG}(\bm{\hat u};\xi_k)=\sum_{i=0}^N\hat u_i\varphi_i(\xi_k)$ which requires $M\cdot Q$ evaluations, the computation of \eqref{eq:numericalFluxIPMBurgers} again requires $O(M\cdot Q)$ operations, i.e. the costs are $O(M^2)$, whereas the evaluation of \eqref{eq:numFluxAnalytic} requires $O(M^3)$ operations.

TODO: Extension to 2D $\rightarrow$ triangular meshes

>>>>

\subsection{Costs of the numerical flux}
A straight forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}, which has been proposed in \cite{kusch2017maximum}. By simply taking moments of a given deterministic flux, the software can easily be extended to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available. However, the need to approximate the flux by an integral seems to be numerically expensive, especially when considering that the stochastic-Galerkin method often allows an analytic computation of all arising integrals. In the following, we discuss this issue and show the advantages of our numerical flux choice by considering the stochastic Burgers' equation
\begin{align*}
\partial_t u + \partial_x \frac{u^2}{2} &= 0,\\
u(t=0,x,\xi) &= u_{IC}(x,\xi).
\end{align*}
The scalar random variable $\xi$ is uniformly distributed in the interval $[-1,1]$, hence the gPC basis functions $\bm\varphi=(\varphi_0,\cdots,\varphi_M)^T$ are the Legendre polynomials. Testing with these basis functions and choosing the SG closure \eqref{eq:SGClosure} yields the closed SG moment system
\begin{align*}
\partial_t \hat u_i + \partial_x \frac12\sum_{n,m = 0}^M \hat u_n \hat u_m \langle \varphi_n\varphi_m\varphi_i \rangle = 0.
\end{align*}
Defining the matrices $\bm C_i := \langle \bm\varphi\bm\varphi^T\varphi_i\rangle\in\mathbb{R}^{M\times M}$ gives
\begin{align*}
\partial_t \bm{\hat u} + \partial_x \bm F(\bm{\hat u}) = 0
\end{align*}
with $F_i(\bm{\hat u}) = \frac12\bm{\hat u}^T\bm C_i\bm{\hat u}$. Note that $\bm{C}_i$ can be computed analytically, hence choosing a Lax-Friedrichs flux
\begin{align}\label{eq:numFluxAnalytic}
G_i^{(LF)}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) =\frac{1}{4}\left(\bm{\hat u}_{\ell}^T \bm{C}_i \bm{\hat u}_{\ell}+\bm{\hat u}_{r}^T \bm{C}_i \bm{\hat u}_{r}\right) - \frac{\Delta x}{2\Delta t}(\bm{\hat u}_{r}-\bm{\hat u}_{\ell})_i
\end{align}
requires no integral evaluations. Recall, that the numerical flux choice made in this work gives
\begin{align}\label{eq:numericalFluxIPMBurgers}
 \bm{G}(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) = \sum_{k=1}^Q w_k g(\mathcal{U}_{SG}(\bm{\hat u}_{\ell};\xi_k),\mathcal{U}_{SG}(\bm{\hat u}_{r};\xi_k))\bm{\varphi}(\xi_k)f_{\Xi}(\xi_k).
\end{align}
When the chosen deterministic flux $g$ is Lax-Friedrichs, the order of the polynomials inside the sum is of order $3M=3(N-1)$. Choosing a Gauss-Lobatto quadrature rule, $Q = \frac32 N -1$ quadrature points suffice for an exact computation of the numerical flux. Indeed, with this choice of quadrature points, the numerical fluxes \eqref{eq:numFluxAnalytic} and \eqref{eq:numericalFluxIPMBurgers} are equivalent. 
Counting the number of operations, one observes that our choice of the numerical flux \eqref{eq:numericalFluxIPMBurgers} is significantly cheaper: When computing and storing the values in a matrix $\bm A\in\mathbb{R}^{Q\times N}$ with entries $a_{ki} = \varphi_i(\xi_k)$ before running the program, the numerical flux \eqref{eq:numericalFluxIPMBurgers} can be split into two parts. First, we determine the SG solution at all quadrature points, i.e. we compute $\bm{u}^{(\ell)} := \bm A \bm{\hat u}_{\ell}$ and $\bm{u}^{(r)} := \bm A \bm{\hat u}_{r}$ which requires $O(N\cdot Q)$ operations. These solution values are then used to compute the numerical flux
\begin{align*}
G_i(\bm{\hat u}_{\ell},\bm{\hat u}_{r}) &= \sum_{k=1}^Q w_k g(u^{(\ell)}_k,u^{(r)}_k)a_{ki}f_{\Xi}(\xi_k),
\end{align*}
which again requires $O(N\cdot Q)$ operations, i.e. the costs are $O(N^2)$. The evaluation of \eqref{eq:numFluxAnalytic} however requires $O(N^3)$ operations.

### 2019-07-11 Paper1/introduction.tex
added adaptive one shot to paper and polished up some parts

When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space. These non-physical structures tend to dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}. Here, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies which locally increase the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or add artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur. The main task here is to find an adequate refinement indicator. \\

In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods. The framework only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers.
Furthermore, we investigate intrusive methods for steady problems and compare them against collocation methods. The steady setting provides different opportunities to take advantage of features of intrusive methods: 
\begin{itemize}
\item Accelerate convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state. 
\item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e. they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}. 
\end{itemize}
In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not). We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.

>>>>

When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space. These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}. Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{durrwachter2018hyperbolicity}. These structures commonly arise on a small portion of the space-time domain. Therefore, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies. By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs. The main task here is to find an adequate refinement indicator. \\

In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods. The framework only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers.
Furthermore, we investigate intrusive methods for steady problems and compare them against collocation methods. The steady setting provides different opportunities to take advantage of features of intrusive methods: 
\begin{itemize}
\item Accelerate the convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state. 
\item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e. they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}. 
\end{itemize}
In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not). We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem.

### 2019-07-11 Paper1/one_shot_ipm.tex
added adaptive one shot to paper and polished up some parts

Since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{(N+1)\cdot N_x\times (N+1)\cdot N_x}$ defined by

@@

The Jacobian $\bm{J}\in\mathbb{R}^{2(N+1)\cdot N_x \times 2(N+1)\cdot N_x}$ has the form

@@

\text{det}(\bm{\tilde J}) = \sum_{\sigma} \text{sgn}(\sigma)\prod_{i = 1}^{2 N_x (N+1)} \tilde J_{\sigma(i),i}.

>>>>

Since we assume that the classical IPM scheme is contractive at its fix point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by

@@

The Jacobian $\bm{J}\in\mathbb{R}^{2N\cdot N_x \times 2N\cdot N_x}$ has the form

@@

\text{det}(\bm{\tilde J}) = \sum_{\sigma} \text{sgn}(\sigma)\prod_{i = 1}^{2 N_x N} \tilde J_{\sigma(i),i}.
